The Relationship Between Gender Identities and Regrades among Undergraduate Students, 2010-2016

### 2024-02-14 outputs/paper/paper.qmd
finalize paper

---
title: "The Relationship Between Gender Identities and Regrades among Undergraduate Students, 2010-2016"
subtitle: "A replication of Li and Zahar;"
author: Faiza Imam, Quang Mai & Catherine Punnoose 
thanks: "Code and data are available at: https://github.com/ponolite/gender_grade.git"
date: today
date-format: long
abstract: "Academia is a complex ecosystem that feeds off a network of social, cultural and innovative influences which often, stem beyond societal norms. This paper replicates the gender implications regarding results in college students and explores potential fluctuations in request patterns over a semester by specifically analyzing the number of classes taken by students and their willingness to submit a grade request throughout the semester. Our results show that there is a trend for female students to submit regrade requests while taking fewer courses compared to a higher course load. Female students experienced grade increases (3%) compared to their male counterparts. We call for further investigation around this topic to understand gender disparities within the world of academia. "
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

@@

# Load data
inst_data <- read.csv(here::here("inputs/data/inst_data_cleaned.csv"))
std_data <- read_csv(here::here("inputs/data/stdsurvey.csv"))

@@

Our replication package was sourced from the openICPSR Portal under the library `readstata13` package [@readStata13]. We retrieved two datasets–‘instsurvey.dta’ and ‘stdsurvey.dta’– and one code file–‘1_FiguresTables.do’–from the data package ‘DataCode’ to examine gender identities of undergraduate students from Colorado State University from 2010 to 2018 and their correlation to regrade requests [@]. Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], `kableExtra` [@rKableExtra], and `knitr` [@rKnitr].

@@

@@

```{r}
#| echo: false
#| message: false
#| label: tbl-genderclass
#| tbl-cap: Cleaned Table of Number of Classes Students Considered for Regrade Requests by Students’ Gender

@@

classes_dta |>
  filter(consider_regrade != "No") |>
  rename("Count" = n, "Number of Classes" = num_class, "Consider Regrade" = consider_regrade)
head(classes_dta, 10) |>
  kable(
    col.names = c("Number of Classes", "Gender", "Consider Regrade", "Count", "Percentage"),
    booktabs = TRUE,
    align = c("c", "c", "c", "c", "c")
  )

@@

# Create a bar graph using ggplot
ggplot(males, aes(x = category, y = value, fill = category)) + ylim(0, 90) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(value)), vjust = -0.5, size = 3) +  # Add labels on top of bars
  labs(x = "Category",
       y = "Percent") + scale_fill_manual(values = c("Grade Increase" = "#4B659D", "No Change" = "#A83B48", "Grade Decrease" = "#B7BA9F")) +  # Adjust colors
  scale_y_continuous(breaks = custom_breaks) +  # Set custom breaks for y-axis
  theme_minimal()

@@

# Create a bar graph using ggplot
ggplot(females, aes(x = category, y = value, fill = category)) + ylim(0, 90) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(value)), vjust = -0.5, size = 3) +  # Add labels on top of bars
  labs(x = "Category",
       y = "Percent") + scale_fill_manual(values = c("Grade Increase" = "#4B659D", "No Change" = "#A83B48", "Grade Decrease" = "#B7BA9F")) +  # Adjust colors
  scale_y_continuous(breaks = custom_breaks) +  # Set custom breaks for y-axis
  theme_minimal()

@@

# Create a bar graph using ggplot
ggplot(males_mt, aes(x = category, y = value, fill = category)) + ylim(0, 90) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(value)), vjust = -0.5, size = 3) +  # Add labels on top of bars
  labs(x = "Category",
       y = "Percent") + scale_fill_manual(values = c("Grade Increase" = "#4B659D", "No Change" = "#A83B48", "Grade Decrease" = "#B7BA9F")) +  # Adjust colors
  scale_y_continuous(breaks = custom_breaks) +  # Set custom breaks for y-axis
  theme_minimal()

@@

# Create a bar graph using ggplot
ggplot(females_mt, aes(x = category, y = value, fill = category)) + ylim(0, 90) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(value)), vjust = -0.5, size = 3) +  # Add labels on top of bars
  labs(x = "Category",
       y = "Percent") + scale_fill_manual(values = c("Grade Increase" = "#4B659D", "No Change" = "#A83B48", "Grade Decrease" = "#B7BA9F")) +  # Adjust colors
  scale_y_continuous(breaks = custom_breaks) +  # Set custom breaks for y-axis
  theme_minimal()

>>>>

---
title: "The Relationship Between Gender Identities and Regrades among Undergraduate Students, 2010-2016"
subtitle: "A replication of Li and Zahar;"
author: Faiza Imam, Quang Mai & Catherine Punnoose 
thanks: "Code and data are available at: https://github.com/ponolite/gender_grade.git, DOI: 10.1257/pol.20210053"
date: today
date-format: long
abstract: "Academia is a complex ecosystem that feeds off a network of social, cultural and innovative influences which often, stem beyond societal norms. This paper replicates the gender implications regarding results in college students and explores potential fluctuations in request patterns over a semester by specifically analyzing the number of classes taken by students and their willingness to submit a grade request throughout the semester. Our results show that there is a trend for female students to submit regrade requests while taking fewer courses compared to a higher course load. Female students experienced grade increases (3%) compared to their male counterparts. We call for further investigation around this topic to understand gender disparities within the world of academia. "
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| echo: false
#| warning: false
#| message: false

@@

# Load data
inst_data <- read.csv(here::here("inputs/data/inst_data_cleaned.csv"))
std_data <- read_csv(here::here("inputs/data/stdsurvey.csv"))
```

@@

Our replication package was sourced from the openICPSR Portal under the library `readstata13` package [@readStata13]. We retrieved two datasets–‘instsurvey.dta’ and ‘stdsurvey.dta’– and one code file–‘1_FiguresTables.do’–from the data package ‘DataCode’ to examine gender identities of undergraduate students from Colorado State University from 2010 to 2018 and their correlation to regrade requests [@GenderDiff]. Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], `kableExtra` [@rKableExtra], and `knitr` [@rKnitr].

@@

```{R}
#| echo: false
#| message: false
#| label: tbl-genderTime
#| tbl-cap: Table of Instructor Survey Regrade Results by Timing and Students’ Gender

# Male Final Grade Data
cleanInstData <- data.frame(
  maleup_avg =  mean(inst_data$maleup_fn, na.rm = TRUE), 
  malesame_avg = mean(inst_data$malesame_fn, na.rm = TRUE), 
  maledown_avg = mean(inst_data$maledown_fn, na.rm = TRUE)
)

cleanInstData |>
  kable(
    col.names = c("Male: Final Grade Increase ", "Male: Final Grade Same ", "Male: Final Grade Decrease ") , 
    align = c("c", "c", "c")
  )

# Female Final Grade Data
cleanInstData <- data.frame(
  femaleup_avg =  mean(inst_data$femaleup_fn, na.rm = TRUE), 
  femalesame_avg = mean(inst_data$femalesame_fn, na.rm = TRUE), 
  femaledown_avg = mean(inst_data$femaledown_fn, na.rm = TRUE)
)

cleanInstData |>
  kable(
    col.names = c("Female: Final Grade Increase ", "Female: Final Grade Same ", "Female: Final Grade Decrease "),
    align = c("c", "c", "c")
  )

# Male During Semester Grade Data
cleanInstData <- data.frame(
  maleup_mt_avg  =  mean(inst_data$maleup_mt, na.rm = TRUE), 
  malesame_mt_avg = mean(inst_data$malesame_mt, na.rm = TRUE), 
  maledown_mt_avg = mean(inst_data$maledown_mt, na.rm = TRUE)
)

cleanInstData |>
  kable(
    col.names = c("Male: During Semester Grade Increase ", "Male: During Semester Grade Same ", "Male: During Semester Grade Decrease ") , 
    align = c("c", "c", "c")
  )

# Female During Semester Grade Data
cleanInstData <- data.frame(
  femaleup_mt_avg  =  mean(inst_data$femaleup_mt, na.rm = TRUE), 
  femalesame_mt_avg = mean(inst_data$femalesame_mt, na.rm = TRUE), 
  femaledown_mt_avg = mean(inst_data$femaledown_mt, na.rm = TRUE)
)

cleanInstData |>
  kable(
    col.names = c("Female: During Semester Grade Increase ", "Female: During Semester Grade Same ", "Female: During Semester Grade Decrease ") , 
    align = c("c", "c", "c")
  )

@@

```{r}
#| echo: false
#| message: false
#| label: tbl-genderclass
#| tbl-cap: Table of Number of Classes Students Considered for Regrade Requests by Students’ Gender

@@

classes_dta |>
  filter(consider_regrade != "No") |>
  rename("Count" = n, "Number of Classes" = num_class, "Consider Regrade" = consider_regrade) |>
  head(12) |>
  kable(
    col.names = c("Number of Classes", "Gender", "Consider Regrade", "Count", "Percentage"),
    booktabs = TRUE,
    align = c("c", "c", "c", "c", "c")
  )

@@

# Create a bar graph using ggplot
ggplot(males, aes(x = category, y = value, fill = category)) + ylim(0, 90) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(value)), vjust = -0.5, size = 3) +  # Add labels on top of bars
  labs(x = "Category",
       y = "Percent") + scale_fill_manual(values = c("Grade Increase" = "#4B659D", "No Change" = "#A83B48", "Grade Decrease" = "#B7BA9F")) +  # Adjust colors
  scale_y_continuous(breaks = custom_breaks) +  # Set custom breaks for y-axis
  theme_minimal() + theme(legend.position = "bottom")

@@

# Create a bar graph using ggplot
ggplot(females, aes(x = category, y = value, fill = category)) + ylim(0, 90) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(value)), vjust = -0.5, size = 3) +  # Add labels on top of bars
  labs(x = "Category",
       y = "Percent") + scale_fill_manual(values = c("Grade Increase" = "#4B659D", "No Change" = "#A83B48", "Grade Decrease" = "#B7BA9F")) +  # Adjust colors
  scale_y_continuous(breaks = custom_breaks) +  # Set custom breaks for y-axis
  theme_minimal() + theme(legend.position = "bottom")

@@

# Create a bar graph using ggplot
ggplot(males_mt, aes(x = category, y = value, fill = category)) + ylim(0, 90) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(value)), vjust = -0.5, size = 3) +  # Add labels on top of bars
  labs(x = "Category",
       y = "Percent") + scale_fill_manual(values = c("Grade Increase" = "#4B659D", "No Change" = "#A83B48", "Grade Decrease" = "#B7BA9F")) +  # Adjust colors
  scale_y_continuous(breaks = custom_breaks) +  # Set custom breaks for y-axis
  theme_minimal() + theme(legend.position = "bottom")

@@

# Create a bar graph using ggplot
ggplot(females_mt, aes(x = category, y = value, fill = category)) + ylim(0, 90) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(value)), vjust = -0.5, size = 3) +  # Add labels on top of bars
  labs(x = "Category",
       y = "Percent") + scale_fill_manual(values = c("Grade Increase" = "#4B659D", "No Change" = "#A83B48", "Grade Decrease" = "#B7BA9F")) +  # Adjust colors
  scale_y_continuous(breaks = custom_breaks) +  # Set custom breaks for y-axis
  theme_minimal() + theme(legend.position = "bottom")
